---
title: "Project 3"
author: "Sasank Desaraju, Phuc Pham, Natalie Geigel, Christopher Ludka"
date: "11/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(car)
library(knitr)
library(emmeans)
library(tidymodels)

```

#Introduction: 

#A Priori: 

# Scientific Hypotheses (Planned Analysis):

```{r planned-data-carpentry}
dataset <- readRDS("./data/group3_data.rds")

tidy_dataset <- dataset %>%
  rename(ID = record_id,
         Flanker_test = tb_flanker_unadj_ss,
         Pic_Seq_Mem_Test = tb_picseq_unadj_ss,
         Moca_Score = moca_unadj,
         Musical_Practice = champs_q17c,
         Age = gi_age,
         Education_Level = gi_degree,
         Second_Language = gi_2ndlang_yn,
         Track_Activity = pac_track_doing,
         Paying_Attention = pac_pay_attention,
         Concentration = pac_concentrate,
         Indecisiveness = bdi_indecisiveness,
         Remembering = pac_remember,
         R_Parahippo_vol = fs_parahpc_vol_r,
         L_Parahippo_vol = fs_parahpc_vol_l,
         Est_Tot_intra_cranial_vol = fs_etiv
         ) %>%
  mutate(Second_Language = if_else(condition = is.na(Second_Language), "No", "Yes")) %>%
  mutate(Musical_Practice = as_factor(Musical_Practice),
         Education_Level = as_factor(Education_Level),
         Second_Language = as_factor(Second_Language),
         Track_Activity = as_factor(Track_Activity),
         Paying_Attention = as_factor(Paying_Attention),
         Concentration = as_factor(Concentration),
         Indecisiveness = as_factor(Indecisiveness),
         Remembering = as_factor(Remembering),
         Moca_Score = as.numeric(Moca_Score),
         Age = as.numeric(Age)
  )

tidy_dataset
```

## Hypothesis 1: The ability to pay attention, proxied by Flanker’s test of attention, can be explained by behavioral habits and a self-reported score when accounting for age and education level. These behavioral habits are ones that have been connected in literature to better brain performance: musical practice and fluency in multiple languages. The self-reported score asks the subject about their ability to pay attention in a distracting environment. 

```{r}
hypo_one_dataset <- tidy_dataset %>%
  select(ID,
         Flanker_test,
         Age,
         Musical_Practice,
         Second_Language,
         Education_Level,
         Track_Activity
  )
  

hypo_one_dataset
```
### Statistical Hypothesis:

Ho: Flanker’s test scores will not be significantly different for different behavioral habits, accounting for age and education. 

Ha: Flanker’s test scores will be significantly different for different behavioral 	habits, accounting for age and education. 

### Statistical Model:
$$\begin{aligned}
\hat{y}_{Flanker.Test} = \beta_{Age}X_{Age}? + \beta_{Music}X_{Music}? +
\beta_{Language}X_{Language}? + \beta_{Education}X_{Education}? + \beta_{Track}X_{Track}?
\end{aligned}$$

### Visualization:
```{r}
#Age vs. Flanker's Test
ggplot(data = hypo_one_dataset, aes(x = Age,
                                    y = Flanker_test)) +
  geom_point() +
  geom_smooth(method = lm, formula = `y` ~ `x`, se = F) 

#Musical Practice vs. Flanker's Test
ggplot(data = hypo_one_dataset, aes(x = Musical_Practice, y = Flanker_test,
                                    color = Musical_Practice)) +
  geom_boxplot() +
  stat_summary(fun.data = mean_se, color = "black")

#Second Language vs. Flanker's Test
ggplot(data = hypo_one_dataset, aes(x = Second_Language, y = Flanker_test,
                                    color = Second_Language)) +
  geom_boxplot() +
  stat_summary(fun.data = mean_se, color = "green") +
  geom_dotplot(binaxis = "y",stackdir = "center", binwidth = 0.8)

#Education Level vs. Flanker's Test
ggplot(data = hypo_one_dataset, aes(x = Education_Level, y = Flanker_test,
                                    color = Education_Level)) +
  geom_boxplot() +
  stat_summary(fun.data = mean_se, color = "black")

#Ability to Pay Attention vs. Flanker's Test
ggplot(data = hypo_one_dataset, aes(x = Track_Activity, y = Flanker_test,
                                    color = Track_Activity)) +
  geom_boxplot() +
  stat_summary(fun.data = mean_se, color = "black")
```
### Model Interpretation and Analysis:
```{r}
hypo1_model <- lm(data = hypo_one_dataset, Flanker_test ~ Age + Musical_Practice + Second_Language + Education_Level)

kable(tidy(hypo1_model), digits = 3)
kable(tidy(Anova(hypo1_model, type = 2, alpha = 0.05)), digits = 3)

```

## Hypothesis 2: The ability to recall from memory, proxied by the NIH Picture Sequence Memory Test, can be explained by the combined effect of relevant behavioral habits, a relevant brain volume, and a self-reported score when accounting for age and education. These behavioral habits are ones that have been connected in literature to better brain performance: musical practice and fluency in multiple languages. The brain volume considered is the parahippocampal volume, known in literature to be important for memory function. The self-reported score asks the subject about their ability to recall items in their day-to-day lives. 

```{r}


```

# Scientific Hypotheses (Planned Exploratory Analysis):
```{r exploratory-data-carpentry}
exp_dataset <- tidy_dataset %>%
  filter(redcap_event_name == "event_1_arm_1") %>%
  select(-redcap_event_name) 
# Y (MoCA Score) ~ fct1 (pac_pay_attention) + fct2 (pac_remember) + fct3 (bdi_indecisiveness) + fct4(pac_concentrate) 



#Model 2: understanding factors as related to cognitive impairment compared to moca scores

ggplot(drop_na(exp_dataset,Paying_Attention, Moca_Score), aes(Paying_Attention,Moca_Score,color = Paying_Attention))+
  geom_boxplot()+
  geom_jitter(alpha=0.5)

ggplot(drop_na(exp_dataset,Remembering, Moca_Score), aes(Remembering,Moca_Score,color = Remembering))+
  geom_boxplot()+
  geom_jitter(alpha=0.5)

ggplot(drop_na(exp_dataset,Indecisiveness, Moca_Score), aes(Indecisiveness,Moca_Score,color = Indecisiveness))+
  geom_boxplot()+
  geom_jitter(alpha=0.5)

ggplot(drop_na(exp_dataset,Concentration, Moca_Score), aes(Concentration,Moca_Score,color = Concentration))+
  geom_boxplot()+
  geom_jitter(alpha=0.5)

# If I drop NA's from every column, the result is only 27 observations which is concerning. It gets rid of indecisive scores of 2
MoCA_dropna <- exp_dataset %>%
  select(Paying_Attention,Remembering,Indecisiveness,Concentration,Moca_Score)  %>%
  drop_na(Paying_Attention,Remembering,Indecisiveness,Concentration)
# Y (MoCA Score) ~ fct1 (pac_pay_attention) + fct2 (pac_remember) + fct3 (bdi_indecisiveness) + fct4(pac_concentrate) 

Mocamod <- lm(MoCA_dropna,
              formula = Moca_Score ~ Paying_Attention + Remembering + Indecisiveness + Concentration)
glance(Mocamod)
tidy(Mocamod)
Anova(Mocamod)
```


```{r volume-basic-visualization}
# Visualize factors and check residuals
# TODO: add outlier checking to basic data

# Sasank
# Model 1: Brain volume model
# Picture Sequence ~  Age + Music + PHCleft + PHCright + Language + Education + Remember

# Musical practice is a number here. SD thinks it makes sense since it plots how much people practice
volume_dataset <- exp_dataset %>%
  rename(Music = Musical_Practice,
         Pic = Pic_Seq_Mem_Test,
         Education = Education_Level,
         Sec_Lang = Second_Language,
         R_PHC = R_Parahippo_vol,
         L_PHC = L_Parahippo_vol,
         Tot_Vol = Est_Tot_intra_cranial_vol) %>%
  mutate(Tot_PHC_normed = (R_PHC + L_PHC) / Tot_Vol) %>%
  mutate(Music = if_else(is.na(Music), 0, as.numeric(Music)))
  

# Age
ggplot(volume_dataset, aes(Age, Pic)) +
  geom_jitter() +
  geom_smooth(method = lm , formula = y ~ x, se = FALSE)

kable(tidy(lm(volume_dataset, formula = Pic ~ Age)),
      caption = 'Age on Memory')


# Music
ggplot(volume_dataset, aes(Music, Pic)) +
  geom_jitter() +
  geom_smooth(method = lm , formula = y ~ x, se = FALSE)

kable(tidy(lm(volume_dataset, formula = Pic ~ Music)),
      caption = 'Musical Practice on Memory')


# PHC volume
ggplot(volume_dataset, aes(Tot_PHC_normed, Pic)) +
  geom_jitter() +
  geom_smooth(method = lm , formula = y ~ x, se = FALSE)

kable(tidy(lm(volume_dataset, formula = Pic ~ Tot_PHC_normed)),
      caption = 'PHC Volume on Memory')


# Second Language
ggplot(volume_dataset, aes(Sec_Lang, Pic)) +
  geom_boxplot() +
  geom_jitter(height = 0)
  #geom_smooth(method = lm , formula = y ~ x, se = FALSE)

kable(tidy(lm(volume_dataset, formula = Pic ~ Sec_Lang)),
      caption = 'Second Language on Memory')


# Education
ggplot(volume_dataset, aes(Education, Pic)) +
  geom_boxplot() +
  geom_jitter(height = 0)
  #geom_smooth(method = lm , formula = y ~ x, se = FALSE)
  
  kable(tidy(lm(volume_dataset, formula = Pic ~ Education)),
      caption = 'Education on Memory')

# With Pic:
# Age is neg corr., Music is not corr., PHC is pos corr., Second_Lang is pos corr., Education is slight pos corr.
```

```{r volume-model-no-ints}
full_vol_model_no_int <- lm(volume_dataset, formula = Pic ~ Age + Music + Tot_PHC_normed + Sec_Lang + Education)

tidy(Anova(full_vol_model_no_int))
glance(full_vol_model_no_int)
```

```{r volume-models-ints}
# Age, PHC, Second Language are significant by themselves. What kind of interactions should we include?
# Maybe second order between all three? Maybe we should do second order between all 5??
# Let's do second order between all three first?

vol_int_1 <- lm(volume_dataset, formula = Pic ~ Age + Music + Tot_PHC_normed + Sec_Lang + Education + Age:Tot_PHC_normed + Tot_PHC_normed:Sec_Lang + Age:Sec_Lang)

tidy(Anova(vol_int_1))
glance(vol_int_1)


vol_int_2 <- lm(volume_dataset, formula = Pic ~ Age*Music*Tot_PHC_normed*Sec_Lang*Education)

full_pic_model_no_int <- lm(volume_dataset, formula = Pic_Seq_Mem_Test ~ Age + Musical_Practice + Tot_PHC_normed + Second_Language + Education_Level)
tidy(Anova(vol_int_2))
glance(vol_int_2)

```

#Unplanned Exploratory Analysis:
```{r}

Moca_unplanned <- exp_dataset %>%
  select(ID,Flanker_test,Pic_Seq_Mem_Test,Moca_Score,Age) 

ggplot(drop_na(Moca_unplanned,Age,Moca_Score), aes(Age,Moca_Score))+
  geom_point()+
  geom_smooth(method = lm, formula = y~x)

ggplot(drop_na(Moca_unplanned,Flanker_test, Moca_Score), aes(Flanker_test,Moca_Score))+
  geom_point()+
  geom_smooth(method = lm, formula = y~x)

ggplot(drop_na(Moca_unplanned,Pic_Seq_Mem_Test, Moca_Score), aes(Pic_Seq_Mem_Test,Moca_Score))+
  geom_point()+
  geom_smooth(method = lm, formula = y~x)

Moca_unplanned_mod <- lm(Moca_unplanned,
                         formula = Moca_Score ~ Age + Flanker_test + Pic_Seq_Mem_Test)
kable(Anova(Moca_unplanned_mod))
kable(tidy(Moca_unplanned_mod))
kable(glance(Moca_unplanned_mod))

```
#Supplemental Data:
